---
title: "Drumkit Project Dev Journal"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

github_sha_link <- function(sha) {
  glue::glue(
    "([changelog: {substr(sha, 1, 6)}]",
    "(https://github.com/gadenbuie/js4shiny-drumkit/commit/{sha}))"
  )
}
```

I've setup a basic [index.html](index.html) to get us started.

# Drumkit: A Single Page App

## Add the drumkit image

`r github_sha_link("195f7b9a1788aeb6afd75803007329c0fea1cbaa")`

Typically images are included in a page using the `<img>` tag.

Let's try that first

```html
<!-- PUT SVG HERE -->
<img src="drumkit.svg" />
```

This doesn't work the way we want.
Compare the [drumkit.svg](drumkit.svg) document
with what we see in the developer console.
The SVG document is XML (like HTML),
but uses a different set of tags.

In order to be able to access the structure of the SVG,
we need to drop the whole SVG into our HTML page.

`r github_sha_link("0faaca62c90b1422a32552620c2a65a6e5835d97")`

## Change the drumkit's colors

`r github_sha_link("4647167e7f1ef0ba83a2bfc38a7f1306af7cfd8c")`

I've include some styles in a `<style>` tag 
at the bottom of the page.
Use these to change the colors of the drumkit
and get a sense for the structure of the svg:

* Individual SVG elements: `polygon`, `path`, `ellipse`
* Each have classes: `.drum-side.kick`
* Are also grouped together, with ID for group: `#kick`
* Note that last class is the same as the sound we want to play: `class="drum-side kick"` means we should use a kick drum sound.

Here are suggestions for colors for the drumkit.

- black stands
- gray rims
- purple drums
- dark purple bottoms
- white kick front
- yellow symbols
- orange hover

Check out the hover effect!

## Try the playDrum() function

In the dev console, try running

```
playDrum('kick')
playDrum('snare')
```

## Find the drumkit and the text elements

`r github_sha_link("8665e8683604070c51e665d672f4ebc5bb411fe9")`

First, find the DOM elements that we're going to want to use.

```js
// Find drumkit, hoverText, and clickText elements
const drumkit = document.getElementById('drumkit');
const hoverText = document.getElementById('hover-text');
const clickText = document.getElementById('click-text');
```

## Listen to mouseover events on the drumkit

`r github_sha_link("4193dcd70f630a50132af42e96fd0afd0d01b245")`

This isn't strictly required to make the drumkit work,
but it gives us a chance to practice working with
browser events.

Add an event listener to the drumkit and output

- The name of the current moused-over tag

- The list of classes on that tag, separated by a dot.

Create the event listener first 
and just write the `event` object to the console. 

```{js}
drumkit.addEventListener('mouseover', function(event) {
  console.log(event)
})
```

Then use `event.target.tagname` and `event.target.classList`.

```{js}
drumkit.addEventListener('mouseover', function(event) {
  const tag = event.target.tagName
  let classes = event.target.classList
  classes = ['', ...classes]
  hoverText.textContent = tag + ' ' + classes.join('.')
})
```

Use the browser dev tools to examine the `<svg>` element.
You should see a small "event" box next to the slement in the DOM.
This tells you that there is an event listener
attached to this element and you can inspect
which events are being watched.

The browser will also highlight in the DOM tree any updates.
Watch the `<p>` element containing the hover text
to see how the browser alerts you to updates in the DOM.

## Add our click event listener

`r github_sha_link("94d5795b1088ff29fb4dc46b57030fa59c355639")`

- Get the `.classList` of the `event.target`.

- Use `.length` to get the last class

- Write out the class to `console.log()`

- Confirm that it's working and then write to the click text element

## Make some noise!

`r github_sha_link("a1d8b45e4daf64ef661b2fd15e82a3f3a9e716dc")`

Store the last class name as `drumPart`...

...and send it off to `playDrum()`.

_insert gif of animal: wipe out!_

## Rewrite event listener to listen on the document

`r github_sha_link("7c8f93e89e9cff37eee7634cf9c635a86e3998a8")`

Instead of listening for click events on `drumkit`,
rewrite the event listener to be attached to the `document`.

Use `element.match('selecter')` to only listen to clicks
on just elements that are

1. direct descendents of a `<g>` element (svg group)

1. which are direct descendents of an `<svg>` element

1. which has the id `.drumkit`.

<details><summary>A convoluted way of saying</summary>

```
svg.drumkit > g > *
```

</details>

You'll know it's working when it makes sound again.

<details><summary>Solution</summary>

```{js}
document.addEventListener('click', function(event) {
  const el = event.target
  if (!el.matches('svg#drumkit > g > *')) {
    // not an event we want to worry about
    return;
  }

  const classes = el.classList
  const drumPart = classes[classes.length -1]
  clickText.textContent = drumPart
  playDrum(drumPart)
})
```
</details>


# A Shiny Drumkit App

We're going to basically repeat the process again,
this time in Shiny.
(And we get to use what we've learned so far.)

## Create a skeleton Shiny app with a module

`r github_sha_link("5392061d0f631f8deec979f1f45eabb80b3eb523")`

1. Create a bare bones Shiny app.
   The `shinyapp` snippet is helpful here.

2. Then make a bare bones Shiny module.
   Again, the `shinymod` snippet is helpful.
   Call the module `drumkitUI` and `drumkit`.

## Embed the drumkit svg using htmlTemplate()

`r github_sha_link("9dafec9246338c5daef405508be4141a1595dee2")`

1. Use `htmltools::htmlTemplate()` to embed the drumkit svg image.
   This is a neat function that lets you embed HTML
   while replacing some of the content dynamically.
   
1. Then add `drumkitUI()` to the page.

1. Run the app. We've got a drumkit, but were' back to square one.

## Build up the drumkitUI() function


Before we go to far, 
lets work out a few details of the `drumkitUI()` function.

### Goal

`r github_sha_link("6ed8b3522f2154fb935bce77730af7171c79f248")`

It will need an ID and we would probably like to be able to set the `width` and `height`.

To do this, we modify the svg to have template variables that can be modified by `htmlTemplate()`.
The template variable needs to appear like `{{ variable }}` in the HTML (or svg) source,
and each variable needs to be on its own line.

### id -> ns("drumkit")

The `id` is going to be the inputId of our drumkit,
and this step is complicated slightly 
by the fact that we're building a module.
We want to be able to reference our `"drumkit"` as `input$drumkit` inside the module.

If we were using regular inputs and Shiny modules, we'd do this:

```{r}
textModuleUI <- function(id, label = "Your Name") {
  ns <- NS(id)
  textInput(ns('name'), label)
}

textModule <- function(input, output, session, ...) {
  observe({
    input$name
  })
}
```

So we need to use the `ID` passed to `drumkitUI()`
to create the module UIs' namespace
and then use that `ns()` function to name our `drumkit` id.

### Finish the template variables

Unfortunately, 
`htmlTemplate()` requires the template variable
to take up the entire line.

This means that we can't do this in our template:

```html
<svg width="{{ width }}" ... />
```

Instead, we need to have `width = "400px"` in the variable passed to the template.
So we have to do a little bit more work.

I'll use the `glue()` package here
becuase I love the package and
it also gives me an opportunity 
to mention how similar `glue()` is
to JavaScript's interpolated string literals.

```{r}
htmlTemplate(
  filename = "drumkit.svg",
  id = glue('id = "{ns("drumkit")}"'),
  width = glue('width = "{width}"'),
  height = glue('height = "{height}"')
)
```

Once you get it working,
try resizing the drumkit and reloading the app.

### Include dependencies

`r github_sha_link("ad2ccaffc19e24c57211a9cf2f9366e63a7ea85e")`

We have the drums now, 
but we need the JavaScript and CSS
we wrote before to make the drums work.

1. Make a folder called `drumkit/`

1. Copy the relevant JavaScript into `drumkit.js`
    - `playDrum()`
    - `document.addEventListener...`
    - update to listen for `.drumkit` class instead of id

1. Copy the CSS into `drumkit.css`

1. Move the `sound/` folder into `drumkit/` too.

!. Might as well move the `drumkit.svg` in there as well.

1. Use `htmltools::htmlDependency()` to "attach" them to the UI function.

   (The `htmldep` snippet can help here too)

Now that dependencies are involved,
we need to stop and start the app to fully reload it.

### Where are the sounds?

`r github_sha_link("40c79f76d2e9adbfc69a6ae9c6772bfa2ae5c84d")`

This almost works but no sound happens.

Check the developer console log to see the error.

<details><summary>Solution</summary>

We need Shiny to serve the `sounds` directory as `sounds/snare.mp3`, etc.

To do this we call on `addResourcePath()`.
</details>

## Connect the Drum Kit to Shiny

The next step is to tell Shiny which drum part was clicked.

### Which drumset?

`r github_sha_link("97c097e85b2eee2a0e03bc61ea0bec411a9293cb")`

Right now we only have one drum set, 
but what if we have multiple drummers on a page?

Inside the event listener, we have two options:

1. Because of our strict filter we know the `<svg>`
   element is the parent of the parent
   of the event target.

1. Or we can use the `.closest()` method
   to find the closest `<svg> element
   
Verify that your method works with `console.log(drumId)`.
   
<details><summary>Solutions</summary>

```{js}
const drumId = el.parentElement.parentElement.id

// or ..

const drumId = el.closest('svg').id
```

</details>

### Report back to Shiny

`r github_sha_link("928d6ceaa04716d367192839b13808bc3841c0a0")`

We'll expore these methods more later in the workshop,
but Shiny provide a method for sending data from JavaScript
back to R.

What we need to know is:

1. The inputId that will receive the value

1. The value to send

We have both of those things already!

```js
Shiny.setInputValue(inputId, value)
```

Then connect the final pieces of the module:

1. Return a reactive from `input$drumkit`

1. call module

1. Add a debug verbatim preview

## Fix the module function

`r github_sha_link("b7b8501ea5c5ac54d16ad5e038d97a9cd9c512cb")`

Move `callModule()` into the module itself.
If you're using the module,
what are the functions inputs that matter most?

## Add a second drummer

`r github_sha_link("680e7e94936c204c818cc0c32210519aec863240")`

To show that it's really really working,
let's add another drummer.

Use `col-xs-6` and set the drumkit size to be responsive (fill container).

Change the drum names (ids) to Mickey and Bill.

## Add a third drummer?
